---
title: "Home Subsidy Algorithm"
author: "Anna Duan"
date: "10/23/2020"
output: html_document
---

---
  title: "Home Repair Tax Credit Program"
author: "Anna Duan"
date: "10/24/2020"
output: 
  html_document: 
  code_folding: hide
toc: true
toc float: true
number_sections: false
---

## Introduction
Motivations for analysis

```{r 1: Develop and interpret data viz describing feature importance and correlation}
```

```{r 2: Split data into 65/35 training/test set}
```

```{r 3: engineer features that increase sensitivity (true positive rate)}
```

```{r 4: interpret features in 1 paragraph}
```

```{r 5: regression summary for kitchen sink & engineered regression}
```

```{r 6: cross validate both models, compare and interpret 2 facet plots of ROC, sensitivity, specificity}
```

```{r 7: ROC curve for new model + interpret}
```

```{r 8: Cost benefit analysis}
```

```{requation for each confusion metric
cost benefit table
plot confusion metric outcomes for each threshold
2 small multiple plots showing threshold as function of total revenue and total count of credits
table of total revenue and total count of credits for 2 categories: 50% threshold and your optimal threshold
Conclusion: should this model be used? How can it be improved? How to make marketing materials result in better response rate?